---
title: "NC Senate Demographics Generate Motion Graphs"
author: "Mark Connolly"
date: "`r Sys.Date()`"
output: html_notebook
---
## Make motion graphs
Generate motion graphs from data and write the results as a series of html files.

Define a collection of district groupings, most of which are grouping of one.  Iterate through the groups and build a motion graph related to race and a second related to ethnicity.  Embed the motion graph in an html document and write the document to a selected file location.  Some of the processing is specific to a Windows environment.

The groupings are somewhat problematic.  The intent is to group districts into a county.  However, districts and county boundaries don't always line up, so the groupings reflect a bit of compromise.  Maybe this data is available geospatially where intersection processing could provide more accurate clustering.  Something to look into.

```{r setup}
# document definition and construction
library(knitr)
library(rmarkdown)

#reading and writing text files for html construction
library(readr)

# data manipulation
library(tidyr)
library(dplyr)

# pipelining commands
library(magrittr)

# dynamic bubble charts
require(googleVis)
```


## Bring in the data

The data used are the data from the data preparation processing.  The data from that step were written to a new csv file.

```{r}
csvfilter <- matrix(c("CSV", "*.csv"), ncol=2)

demosumm <- choose.files(filters = csvfilter,
                    caption = "Select tidy data file",
                    multi=FALSE)  %T>%
           print() %>%
           read_csv()
```

## Set up for processing 

```{r}
work <- data.frame(dist = I(list(1,2,3,4,5,6,7,8,9,10, 11,12,13,
                                 c(14,15,16,17),  # wake county grouping
                                 18,19,20,21,22,23,24,25,26,
                                 c(27,28),  # core guilford
                                 29,30,
                                 c(31,32),  # forsyth
                                 33,34,35,36,
                                 c(37,38,39,40),  # mecklenburg grouping
                                 41,42,43,44,45,46,47,48,49,50
                                )
                            )
                  )
work$race      = I(list(c("white","black","other")))
work$ethnicity = I(list(c("hispanic","nonhisp")))

toname <- function(x,y,z) {
  paste(x, paste(unlist(y), collapse = "_"),z,sep="_")
}

htmlfilter <- matrix(c("HTML", "*.html"), ncol=2)
template <- choose.files(filters = htmlfilter,
                    caption = "Select html template for wrapping motion graph",
                    multi=FALSE)  %T>%
            print() %>%
            read_file()

targetdir <- choose.dir(default = getwd(), 
                        caption = "choose a target to store the generated motion charts")
if (is.na(targetdir)) {
  targetdir <- getwd()
}
print(paste("motion charts will be written to", targetdir))
```

## Define the core function that will process a district to make a motion graph

```{r}
makeamotiongraph <- function(task) {
  
  # by ethnicity
  
  title <- toname("dist",task$dist,"ethnicity")
  
  outfile <- paste(targetdir, "\\", title, ".html", sep="")

  ethnicity <- subset(demosumm,
                      nc_senate %in% unlist(task$dist) & classification %in% unlist(task$ethnicity))

  amotion <-  ethnicity %>%
                gvisMotionChart(idvar = "identity", 
                timevar = "year", 
                xvar = "median_age", 
                yvar = "totalpop_class",
                size = "percent",
                colorvar="nc_senate",
                chartid = title)

  template %>% 
                gsub("{date}", date(), ., fixed=TRUE) %>%
                gsub("{title}", title, ., fixed=TRUE) %>%
                gsub("{motion}", paste((unlist(amotion$html$chart)), collapse=""), ., fixed=TRUE) %>%
                write_file(outfile)

  # by race  
  
  title <- toname("dist",task$dist,"race")
  
  outfile <- paste(targetdir, "\\", title, ".html", sep="")

  race <- subset(demosumm,
                      nc_senate %in% unlist(task$dist) & classification %in% unlist(task$race))

  amotion <-  race %>%
                gvisMotionChart(idvar = "identity", 
                timevar = "year", 
                xvar = "median_age", 
                yvar = "totalpop_class",
                size = "percent",
                colorvar="nc_senate",
                chartid = title)

  template %>% 
                gsub("{date}", date(), ., fixed=TRUE) %>%
                gsub("{title}", title, ., fixed=TRUE) %>%
                gsub("{motion}", paste((unlist(amotion$html$chart)), collapse=""), ., fixed=TRUE) %>%
                write_file(outfile)


}
```
## Process the data
Iterate through the work definitions to produce and persist the motion graphs as shockwave files encapsulated in html documents.
```{r}
for(i in 1:nrow(work)) {
  makeamotiongraph(work[i,])
}

```